Feature importance analysis

### 2020-03-03 analysis/feature-importance.Rmd
wflow_publish(knitr_in("analysis/feature-importance.Rmd"), view = FALSE, verbose = TRUE)

# load drake objects
loadd(
  fi_permut_hr_buffer2,
  fi_permut_vi_buffer2
)
```

Sort the permutation feature importance results

@@

fi_ranked_vi <- fi_permut_vi_buffer2$res %>%
  tibble::rownames_to_column("measure") %>%
  tidyr::pivot_longer(
    cols = starts_with("B"),
    values_to = "importance", names_to = "feature"
  ) %>%
  # mutate(wavelength = seq(420, 995, 4.75)) %>%
  arrange(desc(importance)) %>%
  mutate(rank = row_number()) %>%
  select(-measure)
fi_ranked_vi
```

@@

```{r feature-importance-2 }
spectra_sim <- hsdar::PROSAIL()
spectra_df <- data.frame(
  reflectance = as.vector(spectra_sim@spectra@spectra_ma),
  wavelength = seq(400, 2500, 1)
) %>%
  dplyr::filter(wavelength < 1000) %>%
  # scale the reflectance to fit into the range of the y-axis for the filter
  # ranking (the 10 - is to reverse the scale)
  dplyr::mutate(reflectance = 10 - scale(reflectance,
    center = FALSE,
    scale = max(reflectance, na.rm = TRUE) / 10
  ))

# to be able to plot the sec y axis on the plot, we need to supply the scaled R
# object -> we need to save the object separately
reflectance <- as.vector(spectra_sim@spectra@spectra_ma)
reflectance_scaled <- scale(reflectance,
  center = FALSE,
  scale = max(reflectance, na.rm = TRUE) / 10
)
```

@@

```{r feature-importance-3 }
fi_ranked_hr$wavelength <- round(fi_ranked_hr$wavelength)

data_hr_merged <- left_join(spectra_df, fi_ranked_hr, by = c("wavelength")) %>%
  mutate(class = "HR")

# data_vi_merged <- left_join(spectra_df, fi_ranked_vi, by = c("wavelength")) %>%
#   mutate(class = "vi") %>%
#   mutate(reflectance = as.numeric(.data$reflectance))
```

```{r feature-importance-4, echo=FALSE}
plot_reflectance_imp_rank <- function(data) {
  ggplot(data, aes(x = .data[["wavelength"]], y = .data[["rank"]])) +
    geom_line(aes(x = wavelength, y = reflectance, color = "Spectral Signature of Vegetation"),
      linetype = "solid",
      size = 0.3
    ) +
    geom_point(size = 2, aes(color = "Band"), show.legend = T) +
    scale_x_continuous(limits = c(400, 1000), breaks = scales::pretty_breaks()) +
    scale_y_reverse(
      expand = c(0, 0),
      limits = c(10L, 0),
      # breaks = scales::pretty_breaks(),
      breaks = seq(1, 10, 1),
      sec.axis = sec_axis(~ scale(-.,
        center = FALSE,
        scale = max(., na.rm = TRUE) / -1
      ),
      labels = c(1.0, 0.75, 0.55, 0.25, 0),
      name = "Reflectance [%]"
      )
    ) +
    guides(color = guide_legend(
      title = NULL,
      override.aes = list(
        linetype = c("blank", "solid", "solid"),
        shape = c(19, NA, NA)
      )
    )) +
    labs(
      title = "Permutation-based Variable Importance Results",
      subtitle = paste0(
        "Top ten features of 'VI' and 'HR' datasets along a simulated spectral signature of vegetation.\n",
        "Labels show the hyperspectral band number of dataset 'HR'."
      ),
      caption = "Learner: SVM; 100 Monte-Carlo Iterations",
      y = "Importance Rank", x = "Wavelength [nm]"
    ) +
    ggrepel::geom_label_repel(
      label = data$feature,
      nudge_x = 0.5, nudge_y = 0.5,
      label.size = 0.15
    ) +
    ggpubr::theme_pubclean()
}

@@

@@

### P1 Absolute permutation based Var Imp

```{r fi-permut-hr-abs, dev = c("png", "pdf")}
p1 <- fi_ranked_hr %>%
  plot_reflectance_imp_absolute("numeric_id", class = "HR")
p1
```

### P2 Top ten features along the spectral curve

```{r feature-importance-6}
table_inds <- tibble::tribble(
  ~Rank, ~Name, ~Formula,

  "1", "Vogelmann2", "(R_{734}-R_{747})/(R_{715}+R_{726})",
  "2", "Vogelmann4", "(R_{734}-R_{747})/(R_{715}+R_{720})",
  "3", "Vogelmann", "R_{740}/R_{720}",
  "4", "NPCI", "(R_{680}-R_{430})/(R_{680}+R_{430})",
  "5", "Vogelmann3", "D_{715}/D_{705}",
  # "D2"        , "$D_{705}/D_{722}$",
  # "Datt3"     , "$D_{754}/D_{704}$",
  # "PWI"       , "$R_{900}/R_{970}$",
  # "SR7"       , "$R_{440}/R_{690}$",
  # "SRPI"      , "$R_{430}/R_{680}$"
)
```

```{r fi-permut-vi-hr-ranked, dev = c("png", "pdf"), warning=FALSE}
table_inset <- tibble(x = 0.4, y = 1.15, tb = list(table_inds))

p2 <- data_hr_merged %>%
  plot_reflectance_imp_rank() +
  # Vogelmann 2 (P1)
  geom_segment(aes(
    x = 715, xend = 747,
    y = 1, yend = 1,
    color = "Range of VegIndex"
  ), size = 0.5) +

  # Vogelmann 4 (P2)
  geom_segment(aes(
    x = 720, xend = 747,
    y = 2, yend = 2
  ), color = "green", size = 0.5) +

  # Vogelmann  (P3)
  geom_segment(aes(
    x = 720, xend = 740,
    y = 3, yend = 3
  ), color = "green", size = 0.5) +

  # NPCI (P4)
  geom_segment(aes(
    x = 680, xend = 690,
    y = 4, yend = 4
  ), color = "green", size = 0.5) +
  geom_segment(aes(
    x = 430, xend = 440,
    y = 4, yend = 4
  ), color = "green", size = 0.5) +

  # Vogelmann 3 (P5)
  geom_segment(aes(
    x = 705, xend = 715,
    y = 5, yend = 5,
  ), color = "green", size = 0.5) +

  # D2 (P6)
  geom_segment(aes(
    x = 705, xend = 722,
    y = 6, yend = 6
  ), color = "green", size = 0.5) +

  # Datt3 (P7)
  geom_segment(aes(
    x = 704, xend = 709,
    y = 7, yend = 7
  ), color = "green", size = 0.5) +
  geom_segment(aes(
    x = 754, xend = 759,
    y = 7, yend = 7
  ), color = "green", size = 0.5) +

  # PWI (P8)
  geom_segment(aes(
    x = 900, xend = 910,
    y = 8, yend = 8
  ), color = "green", size = 0.5) +
  geom_segment(aes(
    x = 970, xend = 980,
    y = 8, yend = 8
  ), color = "green", size = 0.5) +

  # SR7 (P9)
  geom_segment(aes(
    x = 440, xend = 450,
    y = 9, yend = 9
  ), color = "green", size = 0.5) +
  geom_segment(aes(
    x = 690, xend = 700,
    y = 9, yend = 9
  ), color = "green", size = 0.5) +

  # SRPI (P10)
  geom_segment(aes(
    x = 430, xend = 440,
    y = 10, yend = 10
  ), color = "green", size = 0.5) +
  geom_segment(aes(
    x = 680, xend = 690,
    y = 10, yend = 10
  ), color = "green", size = 0.5) +
  # annotate(geom = "table", x = 395, y = 0,
  #          label = list(table_inds), size = 2.4,
  #        vjust = 1, hjust = 0)
  ggpmisc::geom_table_npc(
    data = table_inset, aes(npcx = x, npcy = y, label = tb),
    hjust = 1, vjust = 1, size = 2.35
  ) +
  scale_color_manual(
    values = c(
      "Band" = "black", "Spectral Signature of Vegetation" = "black",
      "Range of VegIndex" = "green"
    )
  )

p2
```

Combined

```{r feature-importance-8, fig.width=9.5, fig.asp=1.1}
p1 / p2
```

@@

>>>>

# load drake objects
loadd(
  fi_permut_hr_buffer2,
  fi_permut_vi_buffer2,

  df_wavelengths_from_indices
)
```

Preview the ordered feature importance results for datasets "HR" and "VI".

@@

fi_ranked_vi <- fi_permut_vi_buffer2$res %>%
  tibble::rownames_to_column("measure") %>%
  tidyr::pivot_longer(
    cols = starts_with("B"),
    values_to = "importance", names_to = "feature"
  ) %>%
  # mutate(wavelength = seq(420, 995, 4.75)) %>%
  arrange(desc(importance)) %>%
  mutate(rank = row_number()) %>%
  select(-measure) %>%
  mutate(feature = stringr::str_replace(feature, "bf2_", ""))
fi_ranked_vi
```

@@

```{r feature-importance-2 }
spectra_sim <- hsdar::PROSAIL()
spectra_df <- data.frame(
  reflectance = as.vector(spectra_sim@spectra@spectra_ma),
  wavelength = seq(400, 2500, 1)
) %>%
  dplyr::filter(wavelength < 1000) %>%
  # scale the reflectance to [0, 2] to play nicely with the y-axis later (mean dec in rmse)
  dplyr::mutate(reflectance = scale(reflectance,
    center = FALSE,
    scale = max(reflectance, na.rm = TRUE) / 2
  ))
```

@@

```{r feature-importance-3 }
# round the wavelengths of the HR dataset to match with the simulated ones
fi_ranked_hr$wavelength <- round(fi_ranked_hr$wavelength)

data_hr_merged <- left_join(spectra_df, fi_ranked_hr, by = c("wavelength")) %>%
  mutate(class = "HR") %>%
  mutate(reflectance = as.numeric(.data$reflectance))
```

To label only a subset of the data, a custom data.frame is created.

```{r feature-importance-4 }
df_wavelengths_from_indices_imp <- df_wavelengths_from_indices %>%
  left_join(fi_ranked_vi, by = c("class" = "feature")) %>%
  na.omit()

df_wavelengths_from_indices_label <- df_wavelengths_from_indices_imp %>%
  arrange(rank) %>%
  group_by(class, rank) %>%
  dplyr::slice(1) %>%
  ungroup() %>%
  arrange(rank) %>%
  slice(1:10)

df_wavelengths_from_hr <- data_hr_merged %>%
  arrange(rank) %>%
  group_by(class, rank) %>%
  dplyr::slice(1) %>%
  ungroup() %>%
  arrange(rank) %>%
  slice(1:10) %>%
  select(-class) %>%
  rename(class = feature)

df_label_all <- df_wavelengths_from_indices_label %>%
  bind_rows(df_wavelengths_from_hr)
```

```{r feature-importance-5, echo=FALSE}
plot_reflectance_imp <- function(data) {
  ggplot(data, aes(x = .data[["wavelength"]], y = .data[["importance"]])) +
    geom_line(aes(
      x = .data[["wavelength"]], y = reflectance,
      color = "Spectral Signature of Vegetation"
    ),
    linetype = "solid",
    size = 0.6
    ) +
    # HR: plot only rank 11:last
    geom_point(
      data = data_hr_merged[which(!(data_hr_merged$rank %in% c(1:10))), ],
      aes(color = "HR: 11:122"),
      size = 1.8,
      shape = 5
    ) +
    # HR: plot only rank 1:10
    geom_point(
      data = data_hr_merged[which(data_hr_merged$rank %in% c(1:10)), ],
      aes(color = "HR: 1:10"),
      size = 1.8,
    ) +

    # VI: rank 1:10
    geom_point(
      data = df_wavelengths_from_indices_imp[which((df_wavelengths_from_indices_imp$rank %in% c(1:10))), ],
      aes(x = wavelength, y = importance, group = class),
      size = 2, shape = 4, color = "darkgreen"
    ) +
    # VI: rank 1:10
    geom_line(
      data = df_wavelengths_from_indices_imp[which((df_wavelengths_from_indices_imp$rank %in% c(1:10))), ],
      aes(x = wavelength, y = importance, group = class, color = "VI: 1:10"),
      size = 0.2, linetype = "dashed"
    ) +

    # VI: rank 11:89
    geom_point(
      data = df_wavelengths_from_indices_imp[which(!(df_wavelengths_from_indices_imp$rank %in% c(1:10))), ],
      aes(x = wavelength, y = importance, group = class),
      color = "grey",
      shape = 5, size = 2
    ) +
    # # VI: rank 11:89
    geom_line(
      data = df_wavelengths_from_indices_imp[which(!(df_wavelengths_from_indices_imp$rank %in% c(1:10))), ],
      aes(x = wavelength, y = importance, group = class, color = "VI: 11:89"),
      size = 0.2, linetype = "dashed"
    ) +

    scale_color_manual(values = c(
      "HR: 11:122" = "grey",
      "HR: 1:10" = "black",
      "Spectral Signature of Vegetation" = "black",
      "VI: 1:10" = "darkgreen",
      "VI: 11:89" = "grey"
    )) +
    scale_x_continuous(limits = c(400, 1000), breaks = scales::pretty_breaks()) +
    scale_y_continuous(
      sec.axis = sec_axis(~ scale(-.,
        center = FALSE,
        scale = max(., na.rm = TRUE)
      ),
      labels = c(1.0, 0.75, 0.55, 0.25, 0),
      name = "Scaled Reflectance [0, 1]"
      )
    ) +
    guides(color = guide_legend(
      title = NULL,
      override.aes = list(
        linetype = c("blank", "blank", "solid", "dashed", "dashed"),
        shape = c(19, 5, NA, 4, 4),
        color = c("black", "grey", "black", "darkgreen", "grey")
      )
    )) +
    labs(
      title = "Spatial Variable Importance",
      subtitle = paste0(
        "Top ten features of 'VI' and 'HR' datasets along a simulated spectral signature of vegetation.\n",
        "Labels show the hyperspectral band number for dataset 'HR' and the index name for dataset 'VI'."
      ),
      caption = "Learner: SVM; 100 Monte-Carlo Iterations",
      y = "Mean decrease in RMSE", x = "Wavelength [nm]"
    ) +
    ggrepel::geom_label_repel(
      # only label the best 10 features
      data = df_label_all,
      label = df_label_all$class,
      nudge_x = 0.1, nudge_y = 0.1,
      size = 3
    ) +
    ggpubr::theme_pubclean()
}

@@

# P1 Main plot

```{r fi-permut-vi-hr, dev = c("png", "pdf"), warning=FALSE}
p1 <- data_hr_merged %>%
  plot_reflectance_imp()
p1
```

@@

### P2 Absolute permutation based Var Imp

```{r fi-permut-hr-abs, dev = c("png", "pdf")}
p2 <- fi_ranked_hr %>%
  plot_reflectance_imp_absolute("numeric_id", class = "HR")
p2
```

@@

## Combined

```{r fi-hr-vi-abs-comb, fig.width=9.5, fig.asp=1.1}
p2 / p3
```
